### 2016-03-30 study_workload_1.tex

\section{Study of Parallel Workload~\Rmnum{1}}
%\section{Identify the ``Bully"}
\label{sec:workload-1}

The study of parallel workload~\Rmnum{1} consists of two parts. Firstly, we analyze the over-all network performance when Workload \Rmnum{1} is running under different placement and routing configurations. Secondly, we isolate each application from the workload, analyze its performance  by scrutinizing the traffic going through the routers that serve each job. The analysis allows us to identify the ``bully" in the workload.

\subsection{Network Performance Analysis}
\label{sec: workload-1 network analysis}

\iffalse

\begin{table}[ht]
\begin{center}
\caption{Average communication time by all MPI ranks when Workload I is running on the dragonfly network under six different placement and routing configurations.} 
\label{tab:wkld-commtime}
\begin{tabular}{l c c c c c c }
\toprule % Top horizontal line
\toprule
&\multicolumn{6}{c}{Placement and Routing Configurations} \\ 
\cmidrule(l){2-7}
          & CM & CA & CPA & RM & RA & RPA \\ % Column names row
\midrule % In-table horizontal line
Time(ms)  & 476  & 300  & 301  & 255  & 265  & 264  \\ % Content row 1

\midrule % In-table horizontal line
\bottomrule % Bottom horizontal line
\end{tabular}
\end{center}
\end{table}

The average communication time spent by all MPI ranks when Workload~\Rmnum{1} is running under different placement and routing configurations shown in Table~\ref{tab:wkld-commtime}. The random placement policy performs comparably with all routing policies and outperforms all contiguous placement configurations. The contiguous placement coupled with minimal routing (CM) results in highest communication time. When coupled with (progressive) adaptive routing (CA, CPA), the communication time can be reduced but still higher than random placement polices.

\fi

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/gc-traffic}
        \caption{Global Channel Traffic}
        \label{fig:global-channel-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/lc-traffic}
        \caption{Local Channel Traffic}
        \label{fig:local-channel-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/tl-traffic}
        \caption{Terminal Link Traffic}
        \label{fig:terminal-link-traffic}
    \end{subfigure}\\

    \centering   
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/gc-stime}
        \caption{Global Channel Saturated Time}
        \label{fig:global-channel-stime}
    \end{subfigure}\hfill
     \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/lc-stime}
        \caption{Local Channel Saturated Time}
        \label{fig:local-channel-stime}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/tl-stime}
        \caption{Terminal Link Saturated Time}
        \label{fig:terminal-link-stime}
    \end{subfigure}%
   \caption{The aggregate traffic and saturated time of the dragonfly network when Workload~\Rmnum{1} running under six different placement and routing configurations. ``CA" and ``CPA" performs comparably, the corresponding lines are overlapped. }
   \label{fig:wkld-network-traffic-stime}
\end{figure*}

Figure~\ref{fig:wkld-network-traffic-stime} shows the aggregate traffic for terminal links, local and global channels as well as the corresponding saturated time when Workload~\Rmnum{1} is running on the dragonfly network with six different placement and routing configurations. When contiguous placement coupled with minimal routing(CM) is in use, all the traffic are confined within the contiguously allocated groups, causing congestion on some routers along certain shortest paths from the source to the destination. The congested routers are over-loaded with high volume of traffic going through their local and global channels, shown as the blue dashed line in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}. In the meanwhile, the corresponding saturated time of their local and global channels are also the highest compared with other configurations, shown as the blue dashed line in Figure~\ref{fig:global-channel-stime} and Figure~\ref{fig:local-channel-stime} .

When contiguous placement coupled with progressive adaptive routing (CPA) is in use, the traffic would take non-shortest paths via intermediate routers to void congestion. The congested routers along the shortest paths will not suffer the high volume of traffic as they used to. The traffic through local and global channels are greatly reduced, shown as the green dashed line in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}. The corresponding saturated time on local and global channels also drop a lot, indicating the congestion are alleviated by using progressive adaptive routing. The adaptive routing performs comparably compared with progressive adaptive routing, thus CA is overlapped with CPA in the all the figures above.

Random placement policy performs comparably with three different routing policies.
Random placement can uniformly distribute MPI ranks over the network, balance the traffic load over the network. Shown as the solid lines in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}, no router suffers exceptional high volume of traffic on its local and global channels. When the random placements coupled with minimal routing (RM), the packets can avoid unnecessary intermediate forwarding as introduced by (progressive) adaptive routing, thus generating less traffic. The blue solid line(RM) is lower than the red and green solid line (RA, RPA) in Figure~\ref{fig:global-channel-traffic} and~\ref{fig:local-channel-traffic}.

Due to the same reason, no router suffers exceptionally long saturated time on their global channels when random placement coupled (progressive) adaptive are in use, shown as red and green solid lines in Figure~\ref{fig:global-channel-stime} and~\ref{fig:local-channel-stime}. However, when random placement is couple with minimal routing, the saturated time on certain routers along the shortest paths would increase, shown as the prolonged blue solid line in Figure~\ref{fig:global-channel-stime} and~\ref{fig:local-channel-stime}. Since the minimal routing always chooses the shortest paths, increasing the possibility of congestion along those paths.

%local channel traffic and busy time explanation

%We observe similar phenomena from the local channel traffic and saturated time, shown in Figure \ref{fig:local-channel-traffic}, \ref{fig:local-channel-stime}). Contiguous placement coupled with minimal routing (CM) results in local congestion, and thus the aggregate traffic through local channels of certain routers is quite high. Even with adaptive or progressive adaptive routing, the local congestion caused by grouping all MPI ranks in the same group(s) in contiguous placement can not be eliminated. Random placement is effective at addressing this type of congestion.

%terminal link traffic and busy time explanation
Figure~\ref{fig:terminal-link-traffic} is for the purpose of symmetry, it shows the traffic per rank (aka traffic through terminal links), which stay the same when different placement and routing configurations are in use. All six lines are overlapped in Figure~\ref{fig:terminal-link-traffic}. However, due to the traffic changes on local and channels, the saturated time of terminal links will experience variability when different routing policies are in use, as shown in Figure~\ref{fig:terminal-link-stime}. When contiguous placement coupled with minimal routing(CM), the accumulated congestion on local and global channels cause extremely long saturated time on the terminal links, shown as the blue dashed line in Figure~\ref{fig:terminal-link-stime}. This again manifests the severe congestion caused by using contiguous placement and minimal routing.

\subsection{Individual Application Analysis}
\label{sec: workload-1 app analysis}

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/amg/commtime}
        \caption{AMG}
        \label{fig:amg-commtime}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/mg/commtime}
        \caption{MultiGrid}
        \label{fig:mg-commtime}
    \end{subfigure}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.5 in]{wkld/cr/commtime}
        \caption{CrystalRouter}
        \label{fig:cr-commtime}
    \end{subfigure}%
   \caption{The communication time of all ranks in each application. Random placement and adaptive routing can improve the communication time of MultiGrid and CrystalRouter, while AMG's communication time is greatly prolonged.}
   \label{fig:apps-commtime}
\end{figure*}
\TODO{resume from here}
We isolate each application from Workload~\Rmnum{1} to study the performance application
Figure \ref{fig:apps-commtime} shows the distribution of communication time of three applications in per rank manner, when they running concurrently under 6 different placement and routing configurations listed in Table\ref{tab: placement routing configs}.

As shown in Figure \ref{fig:mg-commtime}, \ref{fig:cr-commtime}, MultiGrid and CrystalRouter suffer the longest communication time when running with contiguous placement and minimal routing (CM). When contiguous placement coupled with adaptive or progressive adaptive routing (CA, CPA), the communication time of both MultiGrid and CrystalRouter are significantly reduced. Random placement further reduces their communication time, and reaches the best performance when coupled with (progressive) adaptive routing (RPA, RA). Both MultiGrid and CrystalRouter show the same trend when they are running under different placement and routing configurations. AMG is quite an exception. The communication time of AMG skyrockets when running under random placement coupled with (progressive) adaptive routing (RPA, RA), shown in Figure \ref{fig:amg-commtime}. The contiguous placement coupled with (progressive) adaptive routing can reach the best performance for AMG. CM is no better than RM, but still much better RPA, RA.

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/lc-traffic}
        \caption{AMG Local Channel Traffic}
        \label{fig:amg-lc-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/lc-traffic}
        \caption{MG Local Channel Traffic}
        \label{fig:mg-lc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/lc-traffic}
        \caption{CR Local Channel Traffic}
        \label{fig:cr-lc-traffic}
    \end{subfigure}\\
    \centering
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/amg/gc-traffic}
        \caption{AMG Global Channel Traffic}
        \label{fig:amg-gc-traffic}
    \end{subfigure}\hfill
    \hspace{1em}%
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/mg/gc-traffic}
        \caption{MG Global Channel Traffic}
        \label{fig:mg-gc-traffic}
    \end{subfigure}\hfill
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[height=1.8 in]{wkld/cr/gc-traffic}
        \caption{CR Global Channel Traffic}
        \label{fig:cr-gc-traffic}
    \end{subfigure}%
   \caption{The aggregate traffic go through the local and global channels of routers serving specific application. ``CA" and ``CPA" perform comparably, the corresponding lines are overlapped. More routers are involved in serving each application when random placement is in use, compared with contiguous placement.}
   \label{fig:3app-lc-gc-traffic}
\end{figure*}

% Show the evidence from network level, explain why AMG is different. 
We look into network level to identify the culprit behind AMG's abnormal behavior with random placement and adaptive routing. By identifying the compute nodes each MPI rank resides on and routers that serving each application, we can see that the traffic go through the routers serving AMG skyrockets when random placement and adaptive routing is in use, shown in Figure \ref{fig:amg-lc-traffic} and \ref{fig:amg-gc-traffic}. Adaptive routing will redirect the traffic of other applications to go through the routers belonging to AMG, and increase the traffic burden on those routers, thus slow down the communication of AMG.

No router will be shared between applications under contiguous placement. A number of successive groups will be allocated to the same application under contiguous placement policy. When contiguous placement coupled with minimal routing, application will be free of interference from its concurrently running peers. However, application communication traffic will be confined within their allocated groups, thus congestion is very likely to happen depending on application communication intensity. As the lest communication-intensive application in Workload \Rmnum{1}, AMG can benefit from contiguous placement. Under contiguous placement, the traffic go through the local and global channels of the routers belonging to AMG is significantly lower than with random placement, shown in Figure \ref{fig:amg-lc-traffic}, \ref{fig:amg-gc-traffic}.

MultiGrid and CrystalRouter are more communication-intensive compared with AMG. They will suffer congestion on both local and global channels when contiguous placement is in use, shown as the dashed line in Figure \ref{fig:mg-lc-traffic}, \ref{fig:mg-gc-traffic} and Figure \ref{fig:cr-lc-traffic}, \ref{fig:cr-gc-traffic}. On the other hand, the random placement can uniformly distribute their traffic all over the network, alleviate the possible congestion, shown as the solid lines in Figure \ref{fig:mg-lc-traffic}, \ref{fig:mg-gc-traffic} and Figure \ref{fig:cr-lc-traffic}, \ref{fig:cr-gc-traffic}.

\subsection{Key Observations}
In summary, based on the results presented in section \ref{sec: workload-1 network analysis} and \ref{sec: workload-1 app analysis}, we have made the following observations.

\emph{Network performance will be significantly improved when random placement and adaptive routing are in use.} Random placement can uniformly distribute MPI ranks of application over the network, and adaptive routing can redirect the traffic from congested routers to other less busy routers. Therefore, the network could reach the status of load-balanced and hot-spots free.

\emph{The performance of less communication-intensive jobs in workload will suffer degradation when network random placement and adaptive routing are in use.} AMG in Workload \Rmnum{1} are ``bullied" by their concurrently running communication-intensive peers, MultiGrid and CrystalRouter. AMG shares routers and groups with MultiGrid and CrystalRouter when random placement is in use. The traffic from MultiGrid and CrystalRouter will be redirected to the routers that serving AMG, slowing down AMG's communication. \footnote{We have tried three different congestion sensing schemes that used in adaptive routing\cite{won-prog-adaptive}, although there are some variations between the results, none of the congestion sensing scheme can prevent the ``bully" from happening.}

\emph{Consistent performance of each application can be guaranteed when contiguous placement and minimal routing are in use.} The router and group sharing between jobs are prohibited in contiguous placement policy. The traffic won't be redirected from congested routers to less busy ones when minimal routing is in use. Thus the interference between concurrently running jobs will be eliminated and the ``bully" behavior will be terminated. 
%However, contiguous placement and minimal routing usually results in local congestion and unbalanced network utilization.